---
fontsize: 12pt
mainfont: Times New Roman
output:
  pdf_document:
    latex_engine: xelatex
  word_document:
    reference_doc: template/RWordTemplate.docx
knit: (function(inputFile, encoding) {
  out_dir <- '../output';
  rmarkdown::render(inputFile, 
  encoding = encoding, 
  output_file = file.path(dirname(inputFile), out_dir, paste0("Activities Report_", Sys.Date(), ".pdf"))) })
params:
  permit: CF-25-014  #update needed in 2028
  year: 2026 #annual update
  iphc_oto_count_EBS: 1333 #annual update iphc_dataservices@iphc.int
  iphc_oto_count_NBS:
  crabmeasurements_count: 39215 #annual update shannon.hennessey@noaa.gov
  crabmeasurements_count_resurvey:

    
---

```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## READ ME:  
## Update year; dates; permit number, if necessary; vessel and vessel length, and crab/stomach/iphc count in YAML
## MikTeXt must be installed and updated (must contact IT for install)

## Crab pathology: check annual crab tech memo (google: afsc crab memo XXXX) # no pathology in 2024

library(knitr)
library(RODBC)
library(dplyr)
library(pander)
library(xtable)
library(rstudioapi)# for Oracle GUI interface
library(tidyverse)
library(janitor)

## make sure you are connected to VPN first (and have Oracle database login)
## if an error occurs, log into Oracle SQL developer and make sure your password is not expired
channel <- odbcConnect(dsn = "AFSC", 
                       uid = rstudioapi::showPrompt(title = "Username", 
                                                    message = "Oracle Username", default = ""), 
                       pwd = rstudioapi::askForPassword("enter password"),
                       believeNRows = FALSE)

# Returns TRUE if user clicks 'Yes', FALSE if 'No'
include_NBS <- rstudioapi::showQuestion(
  title = "Region Selection",
  message = "Do you want to include the NBS in this report?",
  ok = "Yes",
  cancel = "No"
) 

include_crab_resample <- rstudioapi::showQuestion(
  title = "Crab resample?",
  message = "Was there a crab resample this year?",
  ok = "Yes",
  cancel = "No"
) 


```

## Report of `r toString(params$year)` Research Activities
## State of Alaska Fish Resource Permit `r params$permit`
## NOAA Fisheries Alaska Fisheries Science Center – RACE Division

```{r , include=FALSE}

# get unique survey ids to reference later in code
# EBS survey definition id = 98, NBS = 143
survey_id <- sqlQuery(channel, paste0("
                                      select distinct survey_definition_id, survey_id 
                                      from RACE_DATA.V_CRUISES
                                      where year = ", params$year, " and 
                                      survey_definition_id in (98, 143);
                                      "))
survey_id_clean <- survey_id %>%
  janitor::clean_names()

# collapse survey_ids (if multiple, ie: both EBS and NBS) in order to call them inside of an SQL query, specifically in the length section
survey_id_str <- paste(survey_id_clean$survey_id, collapse = ", ")

### stations within 3 miles of AK land (updated 10/17/2024)
AK_waters_stations<- c('F-20', 'H-16')

# get start and end dates of survey
if (include_NBS) {
  # YES: Includes both EBS and NBS
  dates <- sqlQuery(channel, paste0("
    select min(start_time) as start_date,
           max(start_time) as end_date 
    from racebase.haul 
    where cruise = ", params$year, " * 100 + 1 
       OR cruise = ", params$year, " * 100 + 2;
  "))
} else {
  # NO: Includes EBS only
  dates <- sqlQuery(channel, paste0("
    select min(start_time) as start_date,
           max(start_time) as end_date 
    from racebase.haul 
    where cruise = ", params$year, " * 100 + 1;
  "))
}

clean_date <- dates %>%
  janitor::clean_names() %>%
  mutate(
    start_date = as.Date(start_date),
    end_date = as.Date(end_date)
  )

# get vessel names and ID numbers
# works for both EBS and NBS since the same vessels are used
vessel_info <- sqlQuery(channel, paste("
                              select vessel_id, vessel_name from
                              race_data.v_cruises where
                              year = ",params$year," and region = 'BS'
                              ;
                              "))
vessel_info_clean <- vessel_info %>%
  janitor::clean_names() %>%
  mutate(
    vessel_name = str_to_title(vessel_name)
  )

# collapse vessel codes in order to call them inside of an SQL query (specifically in the stomachs section)
vessel_id_str <- paste(vessel_info_clean$vessel_id, collapse = ", ")

## Imports species codes, common and scientific names
# species scientific and common names to join with species_codes later in lengths/otos section
species_names <- sqlQuery(channel, "
                          SELECT species_code, common_name, species_name 
                          
                          FROM RACEBASE.SPECIES;
                          ")
```

# Activity Title:
\raggedright
`r params$year` Eastern Bering Sea Bering Sea Continental Shelf Bottom Trawl Survey of Groundfish and Invertebrate Resources 
\bigskip

# Introduction:
The Resource Assessment and Conservation Engineering (RACE) Division of the National Oceanic and Atmospheric Administration's (NOAA) Alaska Fisheries Science Center (AFSC) conducted the Eastern Bering Sea Continental Shelf Bottom Trawl Survey of Groundfish and Invertebrate Resources from  `r format(clean_date$start_date[1], "%d %B")` to `r format(clean_date$end_date[1], "%d %B, %Y")`. This cruise continued the annual series of eastern Bering Sea (EBS) crab and groundfish stock assessment surveys, which began in 1971. The standard survey covered the Bering Sea shelf between the depths of 20 and 200 m from Bristol Bay to St. Matthew Island. The northern Bering Sea survey, which usually covers the northwestern Bering Sea shelf from Norton Sound to the Bering Strait, is conducted biennially and next scheduled in `r params$year + 1`.

The primary objectives of this survey were to provide the following: 1) data on the distribution, abundance, and biological condition of commercially important groundfish and crab species for the North Pacific Fishery Management Council, 2) catch per unit effort (CPUE) and size composition data for the commercial fisheries of the U.S., and 3) support for ongoing studies on the biology, behavior, and dynamics of key ecosystem components.  

# Methods:
For the standard survey, sampling at survey stations was coordinated between two chartered commercial fishing vessels, the FV *`r vessel_info_clean$vessel_name[1]`* and FV *`r vessel_info_clean$vessel_name[2]`*. Both vessels are house-forward trawlers with stern ramps. Stations were located 20 nm apart and two standard survey stations were located within Alaska state waters (F-20 and H-16). The 26 stations in high-density sampling areas previously located near the Pribilof Islands and south of St. Matthew Island were not included in the standard list of survey stations, as a recent analysis by DeFilippo *et al*. (2023) concluded that removal of these stations would have only minor impacts on the precision and accuracy of groundfish biomass estimates.  Thus, in the interest of maximizing the efficient use of survey resources, sampling was discontinued in the high-density sampling zones around St. Matthew Island and the Pribilof Islands. The current EBS survey design therefore consisted of 350 planned stations, as opposed to the 376 planned stations sampled annually prior to 2024. 

Bottom trawls used for sampling were 83-112 eastern trawls. These nets have a 25.3 m (83 ft) headrope and a 34.1 m (112 ft) footrope. They were towed behind 816 kg, 1.8 X 2.7 m, steel "V"-doors and paired 54.9 m (180.1 ft) dandylines. Each lower dandyline had a 61 cm chain extension connected to the lower wing edge to improve bottom-tending characteristics.  

A digital bathythermograph was attached to the headrope and deployed with each trawl, resulting in oblique depth/temperature profiles of the water column. A bottom contact sensor provided data to assess bottom-tending performance of the net and to determine when the footrope was in contact with the seafloor. Net mensuration sensors were used to assess trawl performance and to provide net geometry data to calculate the area swept by the trawl.  

For additional information on survey methods and protocols, see Markowitz *et al*. (2025).

```{r, echo = F, message = F, warning = F}


## import haul info from Oracle database
## NOTE: removed gear type filter. Filtering takes place later.    

if (include_NBS) {
  # YES: Includes both EBS (98) and NBS (143)
  haul_table <- sqlQuery(channel, paste0("
    SELECT a.cruise, b.year, b.survey_definition_id, a.vessel,
           a.stationid, a.haul, a.performance, a.haul_type, gear, 
           a.gear_depth, a.hauljoin 
    FROM RACEBASE.HAUL a
    JOIN RACE_DATA.V_CRUISES b
      ON a.cruisejoin = b.cruisejoin 
    WHERE survey_definition_id IN (98, 143) 
      AND year = ", params$year, ";
  "))
} else {
  # NO: Includes EBS (98) only
  haul_table <- sqlQuery(channel, paste0("
    SELECT a.cruise, b.year, b.survey_definition_id, a.vessel,
           a.stationid, a.haul, a.performance, a.haul_type, gear, 
           a.gear_depth, a.hauljoin 
           
    FROM RACEBASE.HAUL a
    
    JOIN RACE_DATA.V_CRUISES b
    
      ON a.cruisejoin = b.cruisejoin 
      
      WHERE survey_definition_id = 98 
      
      AND year = ", params$year, ";
  "))
}
```

```{r, echo = F, message = F, warning = F}
################################################################################

## Good/bad hauls for bottom trawl survey for EBS and NBS
##    haul type 3 = standard hauls
##    performance >=0 = hauls within survey standards (good)
##    performance <=0 = hauls outside survey standards (bad) 

EBS_good_hauls <- haul_table %>%
  filter(YEAR == params$year, HAUL_TYPE == 3, PERFORMANCE >= 0, SURVEY_DEFINITION_ID == 98, GEAR == 44) %>%
  tally()

EBS_bad_hauls <- haul_table %>%
  filter(YEAR == params$year, HAUL_TYPE == 3, PERFORMANCE < 0, SURVEY_DEFINITION_ID == 98, GEAR == 44) %>%
  tally() 

# Conditionally calculate NBS hauls
if (include_NBS) {
  NBS_good_hauls <- haul_table %>%
    filter(YEAR == params$year, HAUL_TYPE == 3, PERFORMANCE >= 0, SURVEY_DEFINITION_ID == 143) %>%
    tally()

  NBS_bad_hauls <- haul_table %>%
    filter(YEAR == params$year, HAUL_TYPE == 3, PERFORMANCE < 0, SURVEY_DEFINITION_ID == 143) %>%
    tally()
} else {
  # Set default 0 tallies so downstream code referencing these objects won't break
  NBS_good_hauls <- tibble(n = 0)
  NBS_bad_hauls  <- tibble(n = 0)
}
######################################################

## NOTE: crab resurvey is only conducted in the EBS (survey def ID = 98)
## haul type 17 = crab resurvey

if (include_crab_resample) {
  # YES: Calculate  crab resurvey haul count (HAUL_TYPE == 17)
  resurvey_good_hauls <- haul_table %>%
    filter(YEAR == params$year, HAUL_TYPE == 17, PERFORMANCE >= 0, SURVEY_DEFINITION_ID == 98) %>%
    tally()

  resurvey_bad_hauls <- haul_table %>%
    filter(YEAR == params$year, HAUL_TYPE == 17, PERFORMANCE < 0, SURVEY_DEFINITION_ID == 98) %>%
    tally()
} else {
  # NO: Assign default 0 counts so downstream code relying on these objects won't fail
  resurvey_good_hauls <- tibble(n = 0)
  resurvey_bad_hauls  <- tibble(n = 0)
} 
#######################################################

## Special projects hauls (haul type = 3, bottom trawl; 
##                                   = 15 could be special project or standard
##                                   = 20 (survey modernization project tows)
##                                   = 17 crab resample
##                                   = 0, test tow)
## haul_table only contains gear type 44, which excludes special project (ie: beam trawls)
## More details on special project hauls are located on haul form hard copies

EBS_special_project_hauls <-  haul_table %>%
  filter(YEAR == params$year, HAUL_TYPE != 3, HAUL_TYPE != 17, HAUL_TYPE != 0, 
         PERFORMANCE >= 0, SURVEY_DEFINITION_ID == 98) %>%
  tally()
 
if (include_NBS) {
  NBS_special_project_haul <- haul_table %>%
 filter(YEAR == params$year, HAUL_TYPE != 3, HAUL_TYPE != 17, HAUL_TYPE != 0,
        PERFORMANCE >= 0, SURVEY_DEFINITION_ID == 143) %>%
  tally()
} else {
  # Set default 0 tallies so downstream code referencing these objects won't break
  NBS_special_project_haul <- tibble(n = 0)
}
```

```{r, echo = F, message = F, warning = F}
#########################################################

## Length numbers must come from RACE_DATA because juvenile species_codes are scaled up in RACEBASE

 ## Get basic Bering Sea haul info RACE_DATA.CRUISES
race_data_basichaul <- sqlQuery(channel, paste0("
 SELECT a.cruise_id, cruise, a.survey_id, vessel_id
 
  FROM RACE_DATA.CRUISES a
  
  WHERE survey_id IN (", survey_id_str, ")
  
    AND EXTRACT(YEAR FROM start_date) = ", params$year, ";
    "))

## Takes several minutes to run 
## Haul type 3: Standard bottom sample (preprogrammed station)
##          15: Unknown (additional/extra stations?)
##          20: 15/30 and slope shelf tows (survey modernization projects)
##          17: crab resample
##          edit 2024: add in gear type due to presence of PNE net tows
## Results are not the length data presented to stock assessment (SA)- this code counts all stations, only abundance station = Y are presented to SA
race_data_haul_length <- sqlQuery(channel, paste("
                                  
                                  SELECT cruise_id, haul, haul_type, gear, b.haul_id, species_code, length, frequency
                                  
                                  FROM RACE_DATA.LENGTHS a
                                  
                                  JOIN RACE_DATA.HAULS b
                                  
                                  ON a.haul_id = b.haul_id
                                  
                                  WHERE 
                                  
                                  extract (YEAR FROM b.create_date) = ",params$year," and 
                                  
                                  performance >= 0 
                                    ;
                                  "))

EBS_lengths <- race_data_basichaul %>%
  left_join(race_data_haul_length) %>%
  filter(CRUISE == params$year * 100 + 1) %>%
  tally()

if (include_NBS) {
  NBS_lengths <- race_data_hauls %>%
  left_join(race_data_haul_length) %>%
  filter(CRUISE == params$year * 100 + 2) %>%
  tally()
} else {
  NBS_lengths <- tibble(n = 0)
}

if (include_crab_resample) {
  resurvey_lengths <- BS_lengths %>%
    filter(HAUL_TYPE == 17)
} else {
  resurvey_lengths <- tibble(n = 0)
}
```

```{r, echo = F, message = F, warning = F}
#####################################################################################

## Number of otoliths per species for EBS and overall totals,
## Not filtered by specimen_sample_type- so far no need, all are specimen_sample_type = 1 (last checked 2026)

otoliths <- sqlQuery(channel, paste("
                     SELECT a.species_code, species_name, common_name, age, hauljoin

                     FROM RACEBASE.SPECIMEN a 
                     
                     JOIN RACEBASE.SPECIES b
                     
                     ON a.species_code = b.species_code
                     
                     WHERE
                     cruise = ",params$year," * 100 + 1 OR
                     cruise = ",params$year," * 100 + 2 AND
                     region = 'BS'
                     ;
                     "))

EBS_otos <- otoliths %>%
  dplyr::count(SPECIES_CODE, name = "species_total") 


EBS_halibut_otos <- c(10120, as.numeric(params$iphc_oto_count_EBS)) 

EBS_otos_all_species <- rbind(EBS_otos, EBS_halibut_otos)

EBS_oto_total <- EBS_otos_all_species %>%
  summarise(sum(species_total)) %>%
  rename(oto_total = "sum(species_total)")

# for specimen/voucher table
EBS_otos_all_species$category <- rep("OTOLITH", nrow(EBS_otos_all_species))

# add column with specimen type "otolith" for specimen/voucher table
EBS_otos$category <- rep("OTOLITH", nrow(EBS_otos))

EBS_otos_all_species <- EBS_otos_all_species %>%
  left_join(species_names) %>%
  select(-SPECIES_CODE) %>%
  rename(count = species_total,"Common Name" = COMMON_NAME, "Scientific Name" = SPECIES_NAME)

# filter for stations w/i 3-mile state limit
EBS_otos_filter <- haul_table %>%
  right_join(otoliths) %>%
  filter(CRUISE == params$year * 100 + 1 & STATIONID %in% AK_waters_stations) %>%
  dplyr::count(SPECIES_CODE, name = "species_total") %>%
  left_join(species_names) %>%
  select(-SPECIES_CODE) %>%
  rename(count = species_total,"Common Name" = COMMON_NAME, "Scientific Name" = SPECIES_NAME) 

EBS_otos_filter$category <- rep("OTOLITH", nrow(EBS_otos_filter))

if (include_NBS) {
  
  # Calculate NBS otoliths from haul table and otoliths
  NBS_otos <- haul_table %>%
    right_join(otoliths) %>%
    filter(CRUISE == params$year * 100 + 2) %>%
    count(SPECIES_CODE) %>%
    rename(species_total = "n")

  # Safely handle blank/NULL parameter
  iphc_count <- params$iphc_oto_count_NBS
  
  if (!is.null(iphc_count) && !is.na(iphc_count) && iphc_count != "") {
    NBS_halibut_otos <- c(10120, as.numeric(iphc_count))
    NBS_otos_all_species <- rbind(NBS_otos, NBS_halibut_otos)
  } else {
    NBS_halibut_otos <- NULL
    NBS_otos_all_species <- NBS_otos
  }

  # Process full species table
  NBS_otos_all_species$category <- "OTOLITH"
  NBS_otos_all_species <- NBS_otos_all_species %>%
    left_join(species_names) %>%
    select(-SPECIES_CODE) %>%
    rename(count = species_total, "Common Name" = COMMON_NAME, "Scientific Name" = SPECIES_NAME)

  # Process NBS_otos separately if needed downstream
  NBS_otos$category <- "OTOLITH"
  NBS_otos <- NBS_otos %>%
    left_join(species_names) %>%
    select(-SPECIES_CODE) %>%
    rename(count = species_total, "Common Name" = COMMON_NAME, "Scientific Name" = SPECIES_NAME)

  # Total count across species
  NBS_oto_total <- NBS_otos_all_species %>%
    summarise(oto_total = sum(count, na.rm = TRUE))

} else {
  
  # Default fallback objects for non-NBS years
  NBS_otos              <- tibble()
  NBS_halibut_otos      <- NULL
  NBS_otos_all_species  <- tibble()
  NBS_oto_total         <- tibble(oto_total = 0)
  
}

```

```{r, echo = F, message = F, warning = F}
##########################
  
### OTHER SPECIMEN DATA (STOMACH, PATHOLOGY, CRAB WIDTHS, OTHER SPECIAL PROJECTS)

## CHECK ALL SPECIAL PROJECTS FOR COLLECTIONS! REVIEW YEARLY

EBS_crab_measurements <- params$crabmeasurements_count

if (include_NBS) {  
EBS_resurvey_measurements <-  params$crabmeasurements_count_resurvey
} else {
  EBS_resurvey_measurments <- tibble(n = 0)
}

```

```{r, echo = F, message = F, warning = F}

# stomachs -----
# core species codes always included in stomach reporting
# 21740: walleye Pollock | 21720: Pacific cod | 10110: arrowtooth flounder | 10112: Kamchatka flounder | 10120: Pacific halibut
core_species <- c(21740, 21720, 10110, 10112, 10120) 

collect_stomach_data <- function(region_name, core_codes, species_df, force_prompt = reprompt_stomach_counts) {

  # define cache_file 
  cache_dir  <- "../data"
  cache_file <- file.path(cache_dir, paste0("stomach_cache_", region_name, "_", params$year, ".rds"))
  
  # check if cached data exists for the current year
  # this allows for skipping species code and count entry if it's already been done during the current R session
  if (file.exists(ebs_cache_file)) {
  # If cache exists, ask user if they want to overwrite or keep existing counts
  reprompt_stomach_counts <- rstudioapi::showQuestion(
    title = "Stomach Data Entry",
    message = "Do you want to enter species and counts again?",
    ok = "Yes",
    cancel = "No"
  )
  
  # if user selected "No", load and return cached data immediately
    if (!reprompt) {
      message("--> Loading cached ", region_name, " stomach counts from: ", cache_file)
      return(readRDS(cache_file))
    }
  }
  
  # check if cache exists and if re-prompt = Y (to re-enter data)
  message("--> Launching pop-up prompts for ", region_name, " stomach counts...")
  
  add_codes_str <- rstudioapi::showPrompt(
    title = paste(region_name, "Additional Stomach Species"),
    message = paste0(
      "Enter additional SPECIES_CODEs for ", region_name, " separated by commas (leave blank if none).\n\n",
      "Common species_codes include:\n",
      "Arctic cod 21725\n",
      "saffron cod 21735"
    ),
    default = ""
  )

  additional_codes <- if (!is.null(add_codes_str) && nchar(trimws(add_codes_str)) > 0) {
    as.numeric(trimws(unlist(strsplit(add_codes_str, ","))))
  } else {
    numeric(0)
  }

  all_codes <- unique(c(core_codes, additional_codes[!is.na(additional_codes)]))
  
  target_species <- species_df %>%
    filter(SPECIES_CODE %in% all_codes)

  stomach_list <- list()

  for (i in seq_len(nrow(target_species))) {
    sp_code <- target_species$SPECIES_CODE[i]
    c_name  <- target_species$COMMON_NAME[i]
    s_name  <- target_species$SPECIES_NAME[i]
    
    cnt_input <- rstudioapi::showPrompt(
      title = paste0(region_name, " Stomach Count (", i, "/", nrow(target_species), ")"),
      message = paste0(
        "Enter ", region_name, " stomach count for:\n", 
        c_name, " (Code: ", sp_code, ")\n\n",
        "* Note: Email richard.hibshipman@noaa.gov or jonathan.reum@noaa.gov for stomach species counts"
      ),
      default = "0"
    )
    
    count_val <- as.numeric(cnt_input)
    if (is.na(count_val)) count_val <- 0
    
    stomach_list[[i]] <- tibble(
      "Common Name"     = c_name,
      "Scientific Name" = s_name,
      count             = count_val,
      category          = "STOMACH"
    )
  }
  
  summary_table <- bind_rows(stomach_list)
  total_table   <- summary_table %>% summarise(stomach_total = sum(count, na.rm = TRUE))
  
  # summary and totals in 1 df
  res <- list(summary = summary_table, total = total_table)
  
  # save cache into data/ folder (creates directory if missing)
  if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)
  saveRDS(res, file = cache_file)
  
  return(res)
}

# process EBS data (always present in standard survey scope)
ebs_data            <- collect_stomach_data("EBS", core_species, species_names)
EBS_stomach_summary <- ebs_data$summary
EBS_stomach_total   <- ebs_data$total

# run NBS stomach data if include_NBS = TRUE
if (include_NBS) {
  
  # run interactive prompts for NBS if include_NBS = TRUE
  nbs_data            <- collect_stomach_data("NBS", core_species, species_names)
  NBS_stomach_summary <- nbs_data$summary
  NBS_stomach_total   <- nbs_data$total

} else {
  
  # fallback empty structures to prevent "object not found" errors in downstream report chunks
  NBS_stomach_summary <- tibble()
  NBS_stomach_total   <- tibble(stomach_total = 0)
  
}

```

```{r, echo = F, message = F, warning = F}
#--------------------------------------------
# catch/voucher/otolith data within in 3 mile AK territory waters


EBS_NBS_catch_hauls <- sqlQuery(channel, paste("
                                SELECT a.cruise, species_code, stationid, weight, 
                                       number_fish
                                
                                FROM RACEBASE.CATCH a
                                
                                JOIN RACEBASE.HAUL b
                                
                                ON a.hauljoin = b.hauljoin
                                
                                WHERE 
                                  a.cruise = ",params$year," * 100 + 1 OR
                                  a.cruise = ",params$year," * 100 + 2 AND
                                  a.region = 'BS' AND
                                  performance >= 0 AND 
                                  b.haul_type = 3;
                                  "))

# filter catch data to w/i 3 mile AK stations
within_3mi <- EBS_NBS_catch_hauls %>%
  filter(STATIONID %in% AK_waters_stations) %>%
  group_by(SPECIES_CODE) %>%
  summarise_at(c("WEIGHT", "NUMBER_FISH"), sum, na.rm = TRUE) %>%
  left_join(species_names) %>%
  select(-SPECIES_CODE) %>%
  select(COMMON_NAME, SPECIES_NAME, WEIGHT, NUMBER_FISH) %>%
  arrange(desc(WEIGHT)) %>%
  rename("Common Name" = COMMON_NAME, "Scientific Name" = SPECIES_NAME,
         "Weight (kg)" = WEIGHT, "Number" = NUMBER_FISH)
within_3mi[within_3mi == 0] <- NA

num_fishinvert_count <- within_3mi %>%
  tally()

# get voucher data
vouchers <- sqlQuery(channel, paste("
                     SELECT cruise, hauljoin, common_name, species_name 
                     
                     FROM RACEBASE.CATCH a
                     
                     JOIN RACEBASE.SPECIES b
                     
                     ON a.species_code = b.species_code
                     
                     WHERE
                      region = 'BS' 
                      AND 
                      voucher IS NOT NULL 
                      AND
                      cruise = ",params$year," * 100 + 1 OR
                      cruise = ",params$year," * 100 + 2
                      ;
                     "))


# filter voucher to w/i 3- mile AK stations
vouchers_filter <- vouchers %>%
  left_join(haul_table, by = join_by(HAULJOIN, CRUISE)) %>%
  filter(STATIONID %in% AK_waters_stations)

voucher_count <- vouchers_filter %>%
  tally()



```

# Results:
\raggedright
The FV *`r clean_vessel_info$vessel_name[1]`* and FV *`r clean_vessel_info$vessel_name[2]`* conducted `r EBS_good_hauls + EBS_bad_hauls` bottom trawls during the standard EBS survey. Of those trawls, `r EBS_bad_hauls` were determined to have unsatisfactory performance, resulting in retowing at those stations for acceptable performance. A further `r EBS_special_project_hauls` tows were conducted to support special projects including survey modernization efforts and testing upgraded Marport sensors.


During the standard EBS survey, the scientific crew recorded `r prettyNum(EBS_lengths, big.mark = ",")` randomly selected length measurements by sex from priority fish species. Additionally, `r prettyNum(EBS_crab_widths, big.mark = ",")` crab measurements (lengths/widths) were collected in the EBS and shell conditions were assessed. Sagittal otoliths were extracted from `r prettyNum(EBS_oto_total, big.mark=",")` fish from `r tally(EBS_otos)` targeted species. Correlated length/weight measurements were recorded during otolith collection. A total of `r prettyNum(EBS_stomach_total, big.mark = ",")` fish stomachs from the EBS were extracted and preserved for food habits analysis.


Whole specimens of selected fish and invertebrate species were preserved for use in identification training programs and other research. Various tissue samples were removed and preserved for approved research projects. A total of `r prettyNum(num_fishinvert_count, big.mark=",")` fish and invertebrate taxa were identified in catches within the 3-mile territorial waters (Table 1), and `r prettyNum(voucher_count, big.mark=",")` voucher specimens were collected (Table 2).

\newpage
```{r, results="asis", echo=F, message=F, warning=F}

#pandoc.table(within_3mi, missing = '')
panderOptions("big.mark", ",")
panderOptions("table.alignment.default", "left")
panderOptions("table.split.cells", c(30,30,5,5))
emphasize.italics.cols(2)
#pandoc.table(within_3mi, missing = '')
pander(xtable(within_3mi, caption="EBS Catch Summary Table: \\newline Within 3-mile Territorial Waters", comment=F, caption.placement="top", include.rownames=F))


```

#####

```{r, results="asis", echo=F, message=F, warning=F}

# if else code fills in zeros if no vouchers were collected in AK territory stations
if (nrow(vouchers_filter) > 0) {
EBS_vouchers_count <- vouchers_filter %>%
  mutate_if(is.factor, as.character) %>%
  count(SPECIES_NAME, COMMON_NAME) %>%
  rename(count = n,"Common Name" = COMMON_NAME, "Scientific Name" = SPECIES_NAME ) %>%
  select('Common Name','Scientific Name', count)
} else {
  EBS_vouchers_count <- tibble(
    `Common Name` = "No Vouchers Collected",
    `Scientific Name` = "N/A",
    count = 0
  )
}
```


```{r, results="asis", echo=F, message=F, warning=F}
EBS_vouchers_count$category <- rep("VOUCHER", nrow(EBS_vouchers_count))

EBS_collections <- rbind(EBS_vouchers_count, EBS_otos_filter)

panderOptions("big.mark", ",")
panderOptions("table.alignment.default", "left")
emphasize.italics.cols(2)
pander(xtable(EBS_collections, caption="EBS Voucher and Otolith Collection Specimens: Within 3-mile Territorial Waters", comment=F, caption.placement="top", include.rownames=F))

```

# Citations:

DeFilippo, L., Kotwicki, S., Barnett, L., Richar, J., Litzow, M. A., Stockhausen, W. T., & Palof, K. (2023). Evaluating the impacts of reduced sampling density in a systematic fisheries-independent survey design. *Frontiers in Marine Science*, 10, 1219283.

Markowitz, E. H., Rohan, S. K., Wassermann, S., Charriere, N. E., Anderson, C. B., and Stevenson, D. E. (2025). Results of the 2024 eastern Bering Sea continental shelf bottom trawl survey of groundfish and invertebrate fauna. U.S. Department of Commerce, NOAA Technical Memorandum NMFS-AFSC-499, 203 p.


